# Set up

```{r}
if (!require("xfun")) install.packages("xfun")
pkgs <- c(
  'forecast', 'vars', 'urca', 'MLmetrics', 'yardstick',
  'lubridate', 'tsbox', 'timetk', 'rlist', 'here',
  'extrafont', 'patchwork', 'hrbrthemes', 'ggthemes', 'ggsci', 'scales', 
  'tidyverse', 'vroom', 'readxl', 'writexl', 'tsibble', "tidymodels"
)
xfun::pkg_attach2(pkgs, message = FALSE)

loadfonts(device = "win", quiet = TRUE)
setwd(here())
```

# Load data

```{r}
# rice 
rice_m <- read_csv('data/clean/rice_m.csv')

# rice_price
price <- read_csv("data/raw/rice_price.csv")

# exogenious population and gdp of world
exo <- read_csv("data/raw/wdi_exo.csv")
```

```{r}
# combine
data <- rice_m %>%
  left_join(price, by = "date") %>%
  left_join(exo, by = "date") %>%
  mutate(
    gdppc = na.locf(gdppc),
    pop = na.locf(pop)
  ) %>% as_tsibble(index = date) %>%
  filter(year(date) >= 2003 & year(date) < 2020)

data_ts <- data %>% ts_ts()
data_xts <- data %>% ts_xts()
```

```{r}
# log transformation
ex_hommali <- data_ts[, 'ex_hommali'] %>% log()
ex_white <- data_ts[, 'ex_white'] %>% log()
p_w5_th <- data_ts[, 'p_w5_th'] %>% log()
p_w5_vn <- data_ts[, 'p_w5_vn'] %>% log()
p_h100_th <- data_ts[, 'p_h100_th'] %>% log()
gdppc <- data_ts[, 'gdppc'] %>% log()
gdppc2 <- gdppc**2
pop <- data_ts[, 'pop'] %>% log()

vars_ls <- list(
  ex_hommali = ex_hommali, 
  ex_white = ex_white, 
  p_w5_th = p_w5_th, 
  p_w5_vn = p_w5_vn, 
  p_h100_th = p_h100_th
)
```

```{r}
endo <- ts_c(ex_hommali, ex_white, p_h100_th, p_w5_th)
exo <- ts_c(p_w5_vn, gdppc, gdppc2)
```

```{r}
varx_fit <- VAR(endo, p = 1, exogen = exo, season = 12)
VARselect(endo, exogen = exo, season = 12)
```

```{r}
predict(varx_fit, n.ahead = 12, dumvar = exo[1:12,])
exo[1:12,]
```

# Data Visualization

```{r}
ggplot(rice_m %>% filter(year_th >= 2546), aes(x=date)) +
  geom_line( aes(y=ex_hommali/1e6, color = 'ข้าวหอมมะลิ')) +
  geom_line( aes(y=ex_pathum/1e6, color = 'ข้าวหอมปทุม')) +
  geom_line( aes(y=ex_white/1e6, color = 'ข้าวเจ้าขาว')) +
  geom_line( aes(y=ex_glutinous/1e6, color = 'ข้าวเหนียว')) +
  scale_color_d3() +
  expand_limits(y = 0) +
  scale_y_continuous(label = comma) +
  labs(title = 'ปริมาณส่งออกข้าวสารของไทย',
       x =  "ปี",
       y = 'ล้านตันข้าวสาร'
      ) +
  theme_ipsum(
    base_size = 16,
    base_family = "Athiti Light",
    axis_title_size = 16
  ) +
  theme(
    legend.position = c(1, 1),
    legend.direction = 'horizontal',
    legend.justification = 'right',
    legend.title = element_blank()
  ) +
  scale_x_date(date_breaks = '2 years',
               labels = function(x) year(x)+543
               )

# ggsave('../figures/plot_export_m.png', width = 10, height = 7)
```

# Unit Root Test

```{r}
# get object name as a string
get_objname <- function(x) deparse(substitute(x))

# unit root test for a time series
ur_adf <- function(y, n_diff = 0, varname = NULL, ...) {
  
  varname <- if (is.null(varname)) deparse(substitute(y)) else varname
  y <- if(n_diff == 0) y else diff(y, n_diff)
  
  ur.trend <- ur.df(y, type='trend', lags = 10, selectlags = "BIC", ...)
  ur.drift <- ur.df(y, type='drift', lags = 10, selectlags = "BIC", ...)
  ur.none  <- ur.df(y, type='none' , lags = 10, selectlags = "BIC", ...)

  tstat.trend <- ur.trend@teststat
  tstat.drift <- ur.drift@teststat
  tstat.none  <- ur.none@teststat

  cv.trend <- ur.trend@cval
  cv.drift <- ur.drift@cval
  cv.none  <- ur.none@cval
  
  lags.trend <- dim(ur.trend@testreg$coefficients)[1] - 3
  lags.drift <- dim(ur.drift@testreg$coefficients)[1] - 2
  lags.none  <- dim(ur.none@testreg$coefficients)[1]  - 1

  df_test <- rbind(
    cbind(t(tstat.trend), cv.trend, lags.trend),
    cbind(t(tstat.drift), cv.drift, lags.drift),
    cbind(t(tstat.none) , cv.none , lags.none)
  ) %>% 
    as.data.frame() %>%
    rename(c("lag" = "lags.trend")) %>%
    rownames_to_column("hypo") %>%
    filter(str_starts(hypo, 'tau')) %>%
    mutate(
      result = ifelse(abs(statistic) >= abs(`5pct`), 'Reject', 'Accept'),
      variable = varname,
      level = paste0('d', n_diff),
      star = case_when(
        abs(statistic) > abs(`1pct`) ~ "***",
        abs(statistic) > abs(`5pct`) ~ "** ",
        abs(statistic) > abs(`10pct`) ~ "*  ",
        TRUE ~ "   "
      ),
      hypo = case_when(
        hypo == "tau3" ~ "trend",
        hypo == "tau2" ~ "constant",
        hypo == "tau1" ~ "none"
      ),
      statistic = format(round(statistic, 2), nsmall = 2) %>% 
        as.character %>% str_pad(6, "left")
    ) %>%
    select(variable, hypo, level, everything())
  
  return(df_test)
}

# unit root test in a table format
adf_table <- function(ur_adf) {
  ur_adf %>% 
    mutate(
      stat_star = glue::glue(
        "{statistic}{star}({lag})"
      ) %>% as.character()
    ) %>%
    pivot_wider(
      id_cols = c("variable"),
      names_from = c("level", "hypo"),
      values_from = c("stat_star")
    )
}
```

```{r}
vars_nam <- rep(names(vars_ls), 1)

params <- expand_grid(vars_nam, c(0,1)) %>%
  set_names(c('name', 'ndiff')) %>%
  mutate(ser = vars_ls[name])
```

```{r}
pmap(list(params$ser, params$ndiff, params$name), 
     function(first, second, third) {
       ur_adf(first, second, third)
     }) %>% 
  reduce(rbind) %>% adf_table() #%>%
#   write_xlsx('../results/adf_table.xlsx')
```

# Forecast

Steps:
1. Split training and test set
2. Rolling Forecast
3. Evaluate


- tsf_ets
- tsf_arima
- tsf_var
- tsf_rolling_fcast
- tsf_eval_model
- 


## Define functions


### Timeseries Cross-validate

```{r}
tsf_tscv <- function(data, 
                     initial, 
                     assess = 1, 
                     skip = 0, 
                     cumulative = TRUE) {
  
  n_obs <- length(tk_index(data))
  n_round <- floor(1 + (n_obs - initial - assess) / (skip + 1))
  
  res <- data.frame()
  idx_start <- 1
  idx_end <- initial
  for (i in 1:n_round) {
    res[i, 'id'] <- i
    res[i, 'idx_start'] <- idx_start
    res[i, 'idx_end'] <- idx_end
    idx_start <- ifelse(cumulative == TRUE, 1, idx_start + 1 + skip) 
    idx_end <- idx_end + 1 + skip
    }
  
  subet_ts <- function(data, idx_start, idx_end) {
    data %>% subset(start = idx_start, end = idx_end)
  }
  
  res <- res %>% 
    mutate(
      analysis = pmap(list(idx_start, idx_end), subet_ts, data = data),
      assess = pmap(list(idx_end + 1, idx_end + assess), subet_ts, data = data)
    )
  return(res)
}
```

### Model Evaluation

```{r}
tsf_eval_mod <- function(result, transform = "log") {
  if (transform == "log") {
    inverse_trans <- function(x) exp(x)
  }
  else {
    inverse_trans <- function(x) x
  }
  
  assess <- result$assess %>% unlist() %>% inverse_trans()
  pred <- result$pred %>% unlist() %>% inverse_trans()
  
  res <- data.frame(
      model = result$model[[1]],
      param = result$param[[1]] %>% paste(collapse = "-"),
      rmse = yardstick::rmse_vec(assess, pred) %>% round(),
      mae = yardstick::mae_vec(assess, pred) %>% round(), 
      mape = yardstick::mape_vec(assess, pred) %>% round(2)
    ) %>% 
    mutate(
      assess = list(result$assess),
      pred = list(result$pred)
    )
  return(res)
}
```

### Hyperparameter Tuning

```{r}
tsf_tune_grid <- function(cv, method, params_grid) {
  res <- list()
  n_params <- nrow(params_grid)
  pb <- txtProgressBar(0, n_params, style = 3)
  for (i in 1:n_params) {
    param <- flatten(params_grid[i,])
    tryCatch({
      res[[i]] <- method(cv, param) %>% tsf_eval_mod()
    }, error = function(e){})
    setTxtProgressBar(pb, i)
  }
  return(res %>% reduce(bind_rows))
}
```

### ETS

```{r}
tsf_ets <- function(cv, param) {
  
  fit_ets <- function(y, param) {
    do.call(forecast::ets, 
            list.append(y = y, param))
  }
  
  res <- cv
  
  n_ahead = length(res$assess[[1]])
  
  res <- res %>%
    mutate(
      fitted = map(analysis, fit_ets, param),
      pred = map(fitted, function(x) {
        x %>% forecast(h = n_ahead) %>% as.data.frame() %>% `[[`(1)
        }),
      coef = map(fitted, function(x) x[['par']]),
      param = list(param),
      model = "ets"
      ) %>% 
    select(-analysis, -fitted)
  
  return(res)
}
```

### ARIMA

```{r}
tsf_arima <- function(cv, param) {
  
  fit_arima <- function(y, param) {
    if (param$auto == TRUE) {
      forecast::auto.arima(y) 
    } else {
      param$auto <- NULL
      do.call(forecast::Arima, 
              list.append(y = y, param))
    }
  }
  
  res <- cv
  
  n_ahead = length(res$assess[[1]])
  
  res <- res %>%
    mutate(
      fitted = map(analysis, fit_arima, param),
      pred = map(fitted, function(x) {
        x %>% forecast(h = n_ahead) %>% as.data.frame() %>% `[[`(1)
        }),
      coef = map(fitted, function(x) x[['coef']]),
      param = list(param),
      model = "arima"
      ) %>% 
    select(-analysis, -fitted)
  
  return(res)
}
```

### VAR

```{r}

```

### Test

```{r}
cv_test <- tsf_tscv(ex_hommali, initial = 12*14, assess = 1, skip = 0, cumulative = TRUE)
cv_test[1,]
```

```{r}
params_ets_test <- data.frame(
  model = c("MAA", "MAA"),
  damped = c(FALSE, TRUE)
)
params_ets_test
```

```{r}
tsf_ets(cv_test[1:2,], params_ets_test[1,])
```

```{r}
tsf_ets(cv_test[1:2,], params_ets_test[1,]) %>% tsf_eval_mod()
```

```{r}
tsf_tune_grid(cv_test[1:12,], tsf_ets, params_ets_test)
```

```{r}
params_arima_test <- data.frame(
  model = c("MAA", "MAA"),
  damped = c(FALSE, TRUE)
)
params_ets_test
```

```{r}

```

```{r}

```

## Create TSCV

```{r}
cv_ex_hommali <- tsf_tscv(ex_hommali, initial = 12*14, assess = 2, skip = 0, cumulative = TRUE)
cv_ex_white   <- tsf_tscv(ex_white, initial = 12*14, assess = 2, skip = 0, cumulative = TRUE)
```

## ETS

```{r}
param <- list(model = "MAA", damped = FALSE)

tsf_ets <- function(cv, param) {
  
  fit_ets <- function(y, param) {
    do.call(forecast::ets, 
            list.append(y = y, param))
  }
  
  res <- cv
  
  n_ahead = length(res$assess[[1]])
  
  res <- res %>%
    mutate(
      fitted = map(analysis, fit_ets, param),
      pred = map(fitted, function(x) {
        x %>% forecast(h = n_ahead) %>% as.data.frame() %>% `[[`(1)
        }),
      coef = map(fitted, function(x) x[['par']]),
      param = list(param),
      model = "ets"
      ) %>% 
    select(-analysis, -fitted)
  
  return(res)
}

res <- tsf_ets(cv_ex_hommali, param)

res %>% head()
```

## Evaluate

```{r}
tsf_eval_mod <- function(result, transform = "log") {
  if (transform == "log") {
    inverse_trans <- function(x) exp(x)
  }
  else {
    inverse_trans <- function(x) x
  }
  
  assess <- result$assess %>% unlist() %>% inverse_trans()
  pred <- result$pred %>% unlist() %>% inverse_trans()
  
  res <- data.frame(
      model = result$model[[1]],
      param = result$param[[1]] %>% paste(collapse = "-"),
      rmse = yardstick::rmse_vec(assess, pred) %>% round(),
      mae = yardstick::mae_vec(assess, pred) %>% round(), 
      mape = yardstick::mape_vec(assess, pred) %>% round(2)
    ) %>% 
    mutate(
      assess = list(result$assess),
      pred = list(result$pred)
    )
  return(res)
}

xx <- tsf_eval_mod(res)
```

## ETS Params

```{r}
error_opt <- c("A", "M")
trend_opt <- c("N", "M", "A")
seasonal_opt <- c("N", "A", "M")

params1 <- expand_grid(error_opt, trend_opt, seasonal_opt, FALSE) %>%
  set_names(c('error', "trend", "seasonal", "damped"))

params2 <- expand_grid(error_opt, "A", seasonal_opt, TRUE) %>%
  set_names(c('error', "trend", "seasonal", "damped"))

params_ets <- bind_rows(params1, params2) %>%
  mutate(model = paste0(error, trend, seasonal)) %>%
  select(model, damped)

params_ets[25,] <- list("ZZZ", NA)
params_ets
```

## Tuning

```{r}
params_ets[1,] %>% flatten
```

```{r}
tuning_grid <- function(cv, method, params_grid) {
  res <- list()
  n_params <- nrow(params_grid)
  pb <- txtProgressBar(0, n_params, style = 3)
  for (i in 1:n_params) {
    param <- flatten(param_grid[i,])
    tryCatch({
      res[[i]] <- method(cv, param) %>% eval_mod()
    }, error = function(e){})
    setTxtProgressBar(pb, i)
  }
  return(res %>% reduce(bind_rows))
}
```

```{r}
tune_ets_hommali <- tuning_grid(cv_ex_hommali, tsf_ets, params_ets[1:5,])
tune_ets_white   <- tuning_grid(cv_ex_white, tsf_ets, params_ets[1:5,])
```

```{r}
saveRDS(tune_ets_hommali, "results/tune_ets_hommali.rds")
saveRDS(tune_ets_white, "results/tune_ets_white.rds")

tune_ets_hommali <- readRDS("results/tune_ets_hommali.rds")
tune_ets_white <- readRDS("results/tune_ets_white.rds")
```

```{r}
tune_ets_hommali %>% arrange(rmse)
tune_ets_white %>% arrange(rmse)
```

## ARIMA

```{r}
param <- flatten(params_arima[30, 1:3])

tsf_arima <- function(cv, param) {
  
  fit_arima <- function(y, param) {
    if (param$auto == TRUE) {
      forecast::auto.arima(y) 
    } else {
      param$auto <- NULL
      do.call(forecast::Arima, 
              list.append(y = y, param))
    }
  }
  
  res <- cv
  
  n_ahead = length(res$assess[[1]])
  
  res <- res %>%
    mutate(
      fitted = map(analysis, fit_arima, param),
      pred = map(fitted, function(x) {
        x %>% forecast(h = n_ahead) %>% as.data.frame() %>% `[[`(1)
        }),
      coef = map(fitted, function(x) x[['coef']]),
      param = list(param),
      model = "arima"
      ) %>% 
    select(-analysis, -fitted)
  
  return(res)
}

res <- tsf_arima(cv_ex_hommali, param)

res %>% head()
```

## ARIMA Params

```{r}
p <- 0:2
d <- 0
q <- 0:2
D <- 0
P <- 0:1
Q <- 0:1
auto <- FALSE

params1 <- expand_grid(p, d, q, P, D, Q, auto) %>%
  set_names(c('p', "d", "q", "P", "D", "Q", "auto"))

params_arima <- params1 %>%
  mutate(
    order = pmap(list(p, d, q) , ~c(..1, ..2, ..3)),
    seasonal = pmap(list(P, D, Q) , ~c(..1, ..2, ..3))
    ) %>%
  select(order, seasonal, auto)

params_arima[nrow(params1)+1, 3] <- TRUE
params_arima
```

```{r}

```

```{r}
tune_arima_hommali <- tuning_grid(cv_ex_hommali, tsf_arima, params_arima)
tune_arima_white   <- tuning_grid(cv_ex_white,   tsf_arima, params_arima)
```

```{r}
tune_arima_hommali %>% arrange(-mae)
```

```{r}
# saveRDS(res_arima_hommali, "../results/arima_hommali.rds")
# saveRDS(res_arima_white, "../results/arima_white.rds")

res_arima_hommali <- readRDS("results/arima_hommali.rds")
res_arima_white <- readRDS("results/arima_white.rds")

res_arima_hommali %>% select(-result) %>% arrange(rmse)
res_arima_white %>% select(-result) %>% arrange(rmse)
```

## VAR

```{r}
ex_rice <- ts_c(ex_hommali, ex_white, p_h100_th, p_w5_th, p_w5_vn, gdppc, gdppc2)
```

```{r}
cv_ex_rice <- tsf_tscv_(ex_rice, initial = 12*14, assess = 1, skip = 0, cumulative = TRUE)
```

```{r}
cv_ex_rice$analysis[36]
```

### Hyperparameters

```{r}
type_opt <- c("const", "trend", "both", "none")
season_opt <- c(NA, 12)

params <- expand_grid(type_opt, season_opt) %>%
  set_names(c("type", "seasonal"))
params
```

```{r}
tsf_var <- function(cv, param) {
  
  
  fit_var <- function(y, param) {
    param$season <- ifelse(is.na(param$season), NULL, param$season)
    do.call(forecast::Arima, 
            list.append(y = y, param))
  }
}


tsf_arima <- function(cv, param) {
  
  fit_arima <- function(y, param) {
    if (param$auto == TRUE) {
      forecast::auto.arima(y) 
    } else {
      param$auto <- NULL
      do.call(forecast::Arima, 
              list.append(y = y, param))
    }
  }
  
  res <- cv
  
  n_ahead = length(res$assess[[1]])
  
  res <- res %>%
    mutate(
      fitted = map(analysis, fit_arima, param),
      pred = map(fitted, function(x) {
        x %>% forecast(h = n_ahead) %>% as.data.frame() %>% `[[`(1)
        }),
      coef = map(fitted, function(x) x[['coef']]),
      param = list(param),
      model = "arima"
      ) %>% 
    select(-analysis, -fitted)
  
  return(res)
}
```

```{r}
# endo <- ts_c(ex_hommali, ex_white, p_h100_th, p_w5_th)
# exo <- ts_c(p_w5_vn, gdppc, gdppc2)
ex_rice <- ts_c(ex_hommali, ex_white, p_h100_th, p_w5_th, p_w5_vn, gdppc, gdppc2)
```

```{r}
VARselect(uschange[,1:2], lag.max=8,
  type="const")[["selection"]]
```

```{r}
vars_fcast <- function(data, params, h=1, idx_end = NULL) {
  idx_end <- if(is.null(idx_end)) length(data) else idx_end
  method <- paste0(params, collapse = "-")
  params$y <- data %>% subset(end = idx_end)
  
  if(params$auto == TRUE){
    fitted = auto.arima(params$y) %>% forecast(h=h)
  } else {
    params$auto <- NULL
    fitted = do.call(forecast::Arima, params) %>% forecast(h=h)
  }
  
  res <- data.frame(
    index = idx_end, 
    method = method,
    fitted$model$coef %>% t %>% as.data.frame()
  ) %>%
  mutate(
    yhat = list(fitted %>% as.data.frame() %>% `[[`(1)),
    actual = list(data[(idx_end+1):(idx_end+h)]),
    fitted = list(fitted)
  )
  return(res)
}
```
